Model-based Clustering for Neural Populations

### 2021-09-06 documents/cluster_neural_population.tex
ignore mv-glm, fix that...

The conditional priors for these parameters:
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	Latent vectors \(\left\{ \mathbf{x}_{t} \right\}_{t=1}^T\): the conditional prior is
	defined by
	\[\mathbf{x}_{1} \sim N(\mathbf{x}_{0},\mathbf{Q}_{0})\]
	\[\mathbf{x}_{t + 1}|\mathbf{x}_{t} \sim N(\mathbf{A}\mathbf{x}_{t} + \mathbf{b},\mathbf{Q})\]
	\item
	Initials \(\mathbf{x}_{0}\): assume there are \(J\) clusters,
	\[\mathbf{x}_{0} \sim N(\bm{\mu}_{\mathbf{x}_{00}},\ \mathbf{\Sigma}_{\mathbf{x}_{00}})\]
	, where \(\bm{\mu}_{\mathbf{x}_{00}} = \mathbf{0}_{\text{Jp}}\) and \(\mathbf{\Sigma}_{\mathbf{x}_{00}} = \mathbf{I}_{\text{Jp}}\)
	\item
	Linear mapping (loading) for latent vectors
	\(\left\{ d_{i} \right\}_{i = 1}^{N}\) and
	\(\left\{ \mathbf{c}_{i} \right\}_{i = 1}^{N}\):
	\[\left( d_{i},\mathbf{c}'_{i} \right)' \sim N(\bm{\mu}_{\text{dc}}^{\left( z_{i} \right)},\mathbf{\Sigma}_{\text{dc}}^{(z_{i})})\]
	\item
	Mean and covariance for loading in each cluster
	\(\left\{ \bm{\mu}_{\text{dc}}^{(j)} \right\}_{j}\) and
	\(\left\{ \mathbf{\Sigma}_{\text{dc}}^{(j)} \right\}_{j}\):
	\[\bm{\mu}_{\text{dc}}^{(j)} \sim N(\mathbf{\delta}_{dc0},\mathbf{T}_{dc0})\]
	, where \(\bm{\delta}_{dc0} = \mathbf{0}_{p + 1}\) and
	\(\mathbf{T}_{dc0} = \mathbf{I}_{p + 1}\)

	\[\mathbf{\Sigma}_{\text{dc}}^{(j)} \sim W^{- 1}\left( \Psi_{dc0},\nu_{dc0} \right)\]
	, where \(\nu_{dc0} = p + 1 + 2\) and
	\(\Psi_{dc0} = \mathbf{I}_{p + 1} \times 10^{-4}\)
	\item
	Linear dynamics for latent vectors \(\mathbf{A}\) and \(\mathbf{b}\): if the number of cluster is \(J\)
	\begin{enumerate}
		\def\labelenumi{\alph{enumi}.}
		\item
		No constraints on \(\mathbf{Q}\): denote \(\widetilde{\mathbf{A}} = vec(\mathbf{A})\)
		\[\left( \mathbf{b}',\widetilde{\mathbf{A}} \right)' \sim N\left( \bm{\mu}_{BA_{0}},\mathbf{\Sigma}_{BA_{0}} \right)\]
		, where \(\bm{\mu}_{BA_{0}} = \left( \mathbf{0}_{\text{Jp}}',vec\left( \mathbf{I}_{\text{Jp}} \right)' \right)'\)
		and \(\mathbf{\Sigma}_{BA_{0}} = 0.25\mathbf{I}_{Jp(1 + Jp)}\)
		\item
		Assume \(\mathbf{Q}\) is block-diagonal: denote
		\({\widetilde{\mathbf{A}}}_{j} = vec\left( \mathbf{A}_{j} \right)\)
		\[\left( \mathbf{b}'_{j}, {\widetilde{\mathbf{A}}}'_{j} \right)' \sim N\left( \bm{\mu}_{bA_{0}},\mathbf{\Sigma}_{bA_{0}} \right)\]
		, where \(\bm{\mu}_{bA_{0}} = \left( \mathbf{0}'_{p + (j - 1)p^{2}}, vec\left( \mathbf{I}_{p} \right)', \mathbf{0}_{(J - j)p^{2}}' \right)'\)
		and \(\mathbf{\Sigma}_{bA_{0}} = 0.25\mathbf{I}_{p(1 + pJ)}\)
		\item
		Assume \(\mathbf{Q}\) is diagonal:
		\[\left( b_{k},\mathbf{a}'_{k} \right)' \sim N\left( \mathbf{\mu}_{ba_{k}0},\mathbf{\Sigma}_{ba0} \right)\]
		, where \(\mathbf{\mu}_{ba_{k}0} = \left( 0,\mathbf{e}'_{k} \right)'\) and \(\mathbf{\Sigma}_{ba0} = 0.25\mathbf{I}_{Jp + 1}\)	
	\end{enumerate}
	\item
	Process noise \(\mathbf{Q}\): if the number of cluster is \(J\)
	\begin{enumerate}
		\def\labelenumi{\alph{enumi}.}
		\item
		No constraints on \(\mathbf{Q}\):
		\[\mathbf{Q} \sim W^{- 1}(\Psi_{\mathbf{Q}_{0}},\nu_{\mathbf{Q}_{0}})\]
		, where \(\nu_{\mathbf{Q}_{0}} = Jp + 2\) and
		\(\Psi_{\mathbf{Q}_{0}} = \mathbf{I}_{\text{Jp}} \times 10^{- 4}\). (To make the mean of \(\mathbf{Q}\) loosely centered around
		\(\mathbf{I}_{Jp} \times 10^{- 4}\))
		\item
		Assume \(\mathbf{Q}\) is block-diagonal:
		\[\mathbf{Q}^{(j)} \sim W^{- 1}\left( \Psi_{0},\nu_{0} \right)\]
		, where \(\nu_{0} = p + 2\) and
		\(\Psi_{0} = \mathbf{I}_{p} \times 10^{- 4}\). (To make the mean of \(\mathbf{Q}^{(j)}\) loosely centered around
		\(\mathbf{I}_{p} \times 10^{- 4}\))
		\item
		Assume \(\mathbf{Q}\) is diagonal:
		\[q_{k} \sim IG(\frac{\nu_{0}}{2},\frac{\nu_{0}\sigma_{0}^{2}}{2})\]
		, where \(\nu_{0} = 4\) and \(\sigma_{0}^{2} = 10^{- 4}\).
	\end{enumerate}
\end{enumerate}

@@

\subsubsection{Update \(\left\{ d_{i} \right\}_{i = 1}^{N}\) and
	\(\left\{ \mathbf{c}_{i} \right\}_{i = 1}^{N}\)}
To update efficiently, use Laplace approximation again. Denote
\(\left( d_{i},\mathbf{c}'_{i} \right)' = \bm{\zeta}_{i} \in \mathbb{R}^{p + 1}\)
and \(\left( 1,{\mathbf{x}_{t}'^{\left( z_{i} \right)}} \right) = {{\widetilde{\mathbf{x}}}_{t}'^{\left( z_{i} \right)}}\).
\begin{align*}
	P\left( \bm{\zeta}_{i} \middle| \mathbf{y}_{i},\ \left\{ \mathbf{x}_{t}^{\left( z_{i} \right)} \right\}_{t = 1}^{T},\ldots \right) &= \exp{f\left( \bm{\zeta}_{i} \right)} \approx N\left( \bm{\zeta}_{i}|\bm{\mu}_{\bm{\zeta}_{i}},\mathbf{\Sigma}_{\bm{\zeta}_{i}}\  \right)\\
	\frac{\partial f}{\partial\bm{\zeta}_{i}} &= \frac{\partial l}{\partial\bm{\zeta}_{i}} - {\mathbf{\Sigma}_{\text{dc}}^{\left( z_{i} \right)}}^{- 1}\left( \bm{\zeta}_{i} - \bm{\mu}_{\text{dc}}^{(z_{i})} \right) = \left\lbrack \sum_{t = 1}^{T}{\widetilde{\mathbf{x}}}_{t}^{\left( z_{i} \right)}\left( y_{\text{it}} - \lambda_{\text{it}} \right) \right\rbrack - {\mathbf{\Sigma}_{\text{dc}}^{\left( z_{i} \right)}}^{- 1}\left( \bm{\zeta}_{i} - \bm{\mu}_{\text{dc}}^{(z_{i})} \right)\\
	\frac{\partial^{2}f}{\partial\bm{\zeta}_{i}\partial\bm{\zeta}'_{i}} &= \frac{\partial^{2}l}{\partial\bm{\zeta}_{i}\partial\bm{\zeta}'_{i}} - {\mathbf{\Sigma}_{\text{dc}}^{\left( z_{i} \right)}}^{- 1} = -\left\lbrack \sum_{t = 1}^{T}{\lambda_{\text{it}}{\widetilde{\mathbf{x}}}_{t}^{\left( z_{i} \right)}{{\widetilde{\mathbf{x}}}_{t}'^{\left( z_{i} \right)}}} \right\rbrack - {\mathbf{\Sigma}_{\text{dc}}^{\left( z_{i} \right)}}^{- 1}
\end{align*}
, where \(l\) is Poisson log-likelihood. Then use Newton-Raphson to find
\(\bm{\mu}_{\bm{\zeta}_{i}} = \text{argmax}_{\bm{\zeta}_{i}}\left( f\left( \bm{\zeta}_{i} \right) \right)\)
and \(\mathbf{\Sigma}_{\bm{\zeta}_{i}} = -\left\lbrack \frac{\partial^{2}f}{\partial\bm{\zeta}_{i}\partial\bm{\zeta}'_{i}}\left.  \right|_{\bm{\zeta}_{i} = \bm{\mu}_{\bm{\zeta}_{i}}} \right\rbrack^{-1}\). If initial as the previous step fits fails, simply use prior mean of \(\bm{\mu}_{\bm{\zeta}_{i}}\), i.e. \(\bm{\delta}_{dc0}\).

\subsubsection{Update \(\left\{ \mathbf{\mu}_{\text{dc}}^{(j)} \right\}_{j}\) and \(\left\{ \mathbf{\Sigma}_{\text{dc}}^{(j)} \right\}_{j}\)} \label{loading prior}
As above, denote \(\left( d_{i},\mathbf{c}'_{i} \right)' = \bm{\zeta}_{i} \in \mathbb{R}^{p + 1}\).
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	Mean \(\left\{ \bm{\mu}_{\text{dc}}^{(j)} \right\}_{j}\): by
	conjugacy, \(\bm{\mu}_{\text{dc}}^{(j)} \sim N\left( \bm{\delta}_{\text{dc}},\mathbf{T}_{\text{dc}}\  \right)\)
	\begin{align*}
		\mathbf{T}_{\text{dc}}^{- 1} &= \left( \mathbf{T}_{dc0}^{- 1} + n_{j}{\mathbf{\Sigma}_{\text{dc}}^{(j)}}^{- 1} \right)^{- 1}\\
		\bm{\delta}_{\text{dc}} &= \mathbf{T}_{\text{dc}}\left( \mathbf{T}_{dc0}^{- 1}\bm{\delta}_{dc0} + {\mathbf{\Sigma}_{\text{dc}}^{(j)}}^{- 1}\sum_{i:z_{i} = j}^{}\bm{\zeta}_{i} \right)
	\end{align*}
	\item
	Covariance \(\left\{ \mathbf{\Sigma}_{\text{dc}}^{(j)} \right\}_{j}\): by conjugacy, \(\mathbf{\Sigma}_{\text{dc}}^{(j)} \sim W^{- 1}\left( \Psi_{\text{dc}},\nu_{\text{dc}} \right)\)
	\begin{align*}
		\nu_{\text{dc}} &= n_{j} + \nu_{dc0}\\
		\Psi_{\text{dc}} &= \Psi_{dc0} + \sum_{i:z_{i} = j}^{}{\left( \bm{\zeta}_{i} - \bm{\mu}_{\text{dc}}^{(j)} \right)\left( \bm{\zeta}_{i} - \bm{\mu}_{\text{dc}}^{(j)} \right)'}
	\end{align*}
\end{enumerate}

@@

\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	No constraints on \(\mathbf{Q}\):
	\[P\left( \left( \mathbf{b}',{\widetilde{\mathbf{A}}'} \right)' \middle| \left\{ \mathbf{x}_{t} \right\}_{t = 1}^{T},\mathbf{Q},\ldots \right) \propto \left\lbrack \prod_{t = 2}^{T}{N\left( \mathbf{x}_{t}|\mathbf{A}\mathbf{x}_{t - 1} + \mathbf{b},\mathbf{Q} \right)} \right\rbrack P\left( \left( \mathbf{b}',{\widetilde{\mathbf{A}}'} \right)' \right)\]
	Let
	\(\left( \mathbf{b'},{\widetilde{\mathbf{A}}'} \right)' = \bm{\gamma} \in \mathbb{R}^{Jp(1 + Jp)}\),
	\(\left( 1,\ {\mathbf{x}'_{t - 1}} \right)\bigotimes\mathbf{I}_{\text{Jp}}\mathbf{=}{\widetilde{\mathbf{G}}}_{t - 1} \in \mathbb{R}^{Jp \times Jp(1 + pJ)}\)
	and
	\(\widetilde{\mathbf{G}} = \left( {\widetilde{\mathbf{G}}'}_{1},\ldots,{\widetilde{\mathbf{G}}'}_{T - 1} \right)'\).
	By conjugacy, \(\bm{\gamma}|\left\{ \mathbf{x}_{t} \right\}_{t = 1}^{T},\mathbf{Q},\ldots \sim N(\bm{\mu}_{\bm{\gamma}},\mathbf{\Sigma}_{\bm{\gamma}})\)
	\begin{align*}
		\mathbf{\Sigma}_{\bm{\gamma}} &= \left\lbrack \mathbf{\Sigma}_{BA_{0}}^{- 1} + {\widetilde{\mathbf{G}}}^{\mathbf{'}}\mathbf{I}_{T - 1}\bigotimes\left( \mathbf{Q} \right)^{- 1}\widetilde{\mathbf{G}} \right\rbrack^{- 1}\\
		\bm{\mu}_{\bm{\gamma}} &= \mathbf{\Sigma}_{\bm{\gamma}}\left( \mathbf{\Sigma}_{BA_{0}}^{- 1}\bm{\mu}_{BA_{0}} + {\widetilde{\mathbf{G}}}^{'}\mathbf{I}_{T - 1}\bigotimes\left( \mathbf{Q} \right)^{- 1}\left( \mathbf{x}_{2}^{'},\ldots,\mathbf{x}_{T}^{'} \right)' \right)
	\end{align*}

	\item
	Assume \(\mathbf{Q}\) is block-diagonal:
	\[P\left( \left( \mathbf{b}'_{j}, {\widetilde{\mathbf{A}}'}_{j}\ \right)' \middle| \left\{ \mathbf{x}_{t}^{(j)} \right\}_{t = 1}^{T},\mathbf{Q}^{(j)},\ldots \right) \propto \left\lbrack \prod_{t = 2}^{T}{N\left( \mathbf{x}_{t}^{(j)}|\mathbf{A}_{j}\mathbf{x}_{t - 1} + \mathbf{b}_{j},\mathbf{Q}^{(j)} \right)} \right\rbrack P\left( \left( \mathbf{b}'_{j}, {\widetilde{\mathbf{A}}'}_{j} \right)' \right)\]
	Notice that \(\mathbf{A}_{j}\mathbf{x}_{t - 1} + \mathbf{b}_{j}\mathbf{=}\left( \left( 1,\ {\mathbf{x}'_{t - 1}} \right)\bigotimes\mathbf{I}_{p} \right)\left( \mathbf{b}_{j}^{\mathbf{'}}\mathbf{,}{\widetilde{\mathbf{A}}'}_{j} \right)'\), then the problem is reduced to Bayesian linear regression. Denote \(\left( 1,\ {\mathbf{x}'_{t - 1}} \right)\bigotimes\mathbf{I}_{p}\mathbf{=}{\widetilde{\mathbf{X}}}_{t - 1} \in \mathbb{R}^{p \times p(1 + pJ)}\),
	\(\widetilde{\mathbf{X}} = \left({\widetilde{\mathbf{X}}'}_{1},\ldots,{\widetilde{\mathbf{X}}'}_{T - 1} \right)'\)
	and\(\left( \mathbf{b}'_{j}, {\widetilde{\mathbf{A}}'}_{j} \right)' = \mathbf{\gamma}_{j} \in \mathbb{R}^{p(1 + pJ)}\).
	Then
	\begin{align*}
		\prod_{t = 2}^{T}{N\left( \mathbf{x}_{t}^{(j)}|\mathbf{A}_{j}\mathbf{x}_{t - 1} + \mathbf{b}_{j},\mathbf{Q}^{(j)} \right)} &= \prod_{t = 2}^{T}{N\left( \mathbf{x}_{t}^{(j)}|{\widetilde{\mathbf{X}}}_{t - 1}\mathbf{\gamma}_{j},\mathbf{Q}^{(j)} \right)}\\
		&= N\left( \left( {\mathbf{x}_{2}'^{(j)}},\ldots,{\mathbf{x}_{T}'^{(j)}} \right)'|\widetilde{\mathbf{X}}\bm{\gamma}_{j},\ \mathbf{I}_{T - 1}\bigotimes\mathbf{Q}^{(j)} \right)
	\end{align*}
	By conjugacy, \(P\left( \bm{\gamma}_{j} \middle| \left\{ \mathbf{x}_{t}^{(j)} \right\}_{t = 1}^{T},\mathbf{Q}^{(j)},\ldots \right) = N\left(\bm{\gamma}_{j}|\bm{\mu}_{\bm{\gamma}_{j}},\mathbf{\Sigma}_{\bm{\gamma}_{j}} \right)\)
	\begin{align*}
		\mathbf{\Sigma}_{\bm{\gamma}_{j}} &= \left\lbrack \mathbf{\Sigma}_{bA_{0}}^{- 1} + {\widetilde{\mathbf{X}}'}\mathbf{I}_{T - 1}\bigotimes\left( \mathbf{Q}^{(j)} \right)^{- 1}\widetilde{\mathbf{X}} \right\rbrack^{- 1}\\
		\bm{\mu}_{\bm{\gamma}_{j}} &= \mathbf{\Sigma}_{\bm{\gamma}_{j}}\left( \mathbf{\Sigma}_{bA_{0}}^{- 1}\bm{\mu}_{bA_{0}} + {\widetilde{\mathbf{X}}'}\mathbf{I}_{T - 1}\bigotimes\left( \mathbf{Q}^{(j)} \right)^{- 1}\left( {\mathbf{x}_{2}'^{(j)}},\ldots,{\mathbf{x}_{T}'^{(j)}} \right)' \right)
	\end{align*}

	\item
	Assume \(\mathbf{Q}\) is diagonal:
	\[P\left( \left( b_{k},\mathbf{a}'_{k} \right)'|\left\{ x_{kt} \right\}_{t = 1}^{T},q_{k},\ \ldots \right) \propto \left\lbrack \prod_{t = 2}^{T}{N\left( x_{\text{kt}}|\mathbf{a}'_{k}\mathbf{x}_{t - 1} + b_{k},q_{k} \right)} \right\rbrack P\left( \left( b_{k},\mathbf{a}'_{k} \right)' \right)\]
	Rewrite \(\mathbf{a}'_{k}\mathbf{x}_{t - 1} + b_{k} = \left( 1,\mathbf{x}'_{t - 1} \right)\left( b_{k},\mathbf{a}'_{k} \right)'\). Then the problem is reduced to Bayesian linear regression. Denote \(\left( b_{k},\mathbf{a}'_{k} \right)' = \bm{\gamma}_{k} \in \mathbb{R}^{Jp + 1}\)
	and
	\(\widetilde{\mathbf{X}} = \left( \left( 1,\mathbf{x}'_{1} \right)',\ldots,\left( 1,\mathbf{x}'_{T - 1} \right)' \right)^{'}\). Then by conjugacy, \(P\left( \bm{\gamma}_{k}|\left\{ x_{\text{kt}} \right\}_{t = 1}^{T},q_{k},\ \ldots \right) = N\left( \bm{\gamma}_{k}|\bm{\mu}_{\bm{\gamma}_{k}},\mathbf{\Sigma}_{\bm{\gamma}_{k}} \right)\)
	\begin{align*}
		\mathbf{\Sigma}_{\bm{\gamma}_{k}} &= \left\lbrack \mathbf{\Sigma}_{ba0}^{- 1} + \frac{{\widetilde{\mathbf{X}}'}\widetilde{\mathbf{X}}}{q_{k}} \right\rbrack^{- 1}\\
		\bm{\mu}_{\bm{\gamma}_{k}} &= \mathbf{\Sigma}_{\mathbf{\gamma}_{k}}\left( \mathbf{\Sigma}_{ba0}^{- 1}\mathbf{\mu}_{ba0} + {\widetilde{\mathbf{X}}'}\left( x_{k2},\ldots,x_{\text{kT}} \right)'/q_{k} \right)
	\end{align*}	
\end{enumerate}

\subsubsection{Update \(\mathbf{Q}\)}
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	No constraints on \(\mathbf{Q}\):\\
	Let \(\mathbf{A}\mathbf{x}_{t - 1} + \mathbf{b} = \bm{\mu}_{x_{t}}\)
	\[P\left( \mathbf{Q} \middle| \left\{ \mathbf{x}_{t} \right\}_{t = 1}^{T},\mathbf{A},\mathbf{b},\ldots \right) \propto \left\lbrack \prod_{t = 2}^{T}{N\left( \mathbf{x}_{t}|\bm{\mu}_{\mathbf{x}_{t}},\ \mathbf{Q} \right)} \right\rbrack W^{- 1}(\mathbf{Q}|\Psi_{\mathbf{Q}_{0}},\nu_{\mathbf{Q}_{0}})\]
	By conjugacy,
	\[P\left( \mathbf{Q} \middle| \left\{ \mathbf{x}_{t}^{(j)} \right\}_{t = 1}^{T},\ \mathbf{A},\mathbf{b},\ldots \right) = W^{- 1}\left( \mathbf{Q}|\Psi_{0} + \sum_{t = 2}^{T}{\left( \mathbf{x}_{t} - \bm{\mu}_{\mathbf{x}_{t}} \right)\left( \mathbf{x}_{t} - \bm{\mu}_{\mathbf{x}_{t}} \right)^{'}},\ T - 1 + \nu_{0} \right)\]

	\item
	Assume \(\mathbf{Q}\) is block-diagonal:\\
	Let \(\mathbf{A}_{j}\mathbf{x}_{t - 1} + \mathbf{b}_{j}\mathbf{=}\bm{\mu}_{\mathbf{x}_{t}^{(j)}}\),
	\[P\left( \mathbf{Q}^{(j)} \middle| \left\{ \mathbf{x}_{t}^{(j)} \right\}_{t = 1}^{T},\ \mathbf{A}_{j},\mathbf{b}_{j},\ldots \right) \propto \left\lbrack \prod_{t = 2}^{T}{N\left( \mathbf{x}_{t}^{(j)}|\bm{\mu}_{\mathbf{x}_{t}^{(j)}},\mathbf{Q}^{(j)} \right)} \right\rbrack W^{- 1}\left( \mathbf{Q}^{(j)}|\Psi_{0},\nu_{0} \right)\]
	By conjugacy,
	\[P\left( \mathbf{Q}^{(j)} \middle| \left\{ \mathbf{x}_{t}^{(j)} \right\}_{t = 1}^{T},\ \mathbf{A}_{j},\mathbf{b}_{j},\ldots \right) = W^{- 1}\left( \mathbf{Q}^{(j)}|\Psi_{0} + \sum_{t = 2}^{T}{\left( \mathbf{x}_{t}^{(j)} - \bm{\mu}_{\mathbf{x}_{t}^{(j)}} \right)\left( \mathbf{x}_{t}^{(j)} - \bm{\mu}_{\mathbf{x}_{t}^{(j)}} \right)^{'}},\ T - 1 + \nu_{0} \right)\]

	\item
	Assume \(\mathbf{Q}\) is diagonal:\\
	Let \(\mathbf{a}_{k}^{'}\mathbf{x}_{t - 1} + b_{k} = \mu_{x_{\text{kt}}}\),
	\[P\left( q_{k}|\left\{ x_{\text{kt}} \right\}_{t = 1}^{T},b_{k},\mathbf{a}_{k}^{'},\ \ldots \right) \propto \left\lbrack \prod_{t = 2}^{T}{N\left( x_{\text{kt}}|\mu_{x_{\text{kt}}},q_{k} \right)} \right\rbrack\text{IG}\left( q_{k} \middle| \frac{\nu_{0}}{2},\frac{\nu_{0}\sigma_{0}^{2}}{2} \right)\]
	By conjugacy,
	\[P\left( q_{k}|\left\{ x_{\text{kt}} \right\}_{t = 1}^{T},b_{k},\mathbf{a}_{k}^{'},\ \ldots \right) = IG\left( q_{k}|\frac{\nu_{0} + T - 1}{2},\frac{\nu_{0}\sigma_{0}^{2} + \sum_{t = 2}^{T}\left( x_{\text{kt}} - \mu_{x_{\text{kt}}} \right)^{2}}{2} \right)\]
\end{enumerate}

\section{Simulations}
\textcolor{red}{\textbf{Currently, the latent \(x_t\) estimation is not accurate enough. This may come from bad simulation/ slow convergence or other sources. See details in section \ref{problems}}}\\

@@

\clearpage
\section{TODO}
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	Use a more identifiable simulation, e.g. larger loading, to see the performance
	\item
	Currently, the mixing of \(\mathbf{b}\) and \(\mathbf{A}\) is terrible, and I need find a way to break the dependence between dynamics (\(\mathbf{b}\) and \(\mathbf{A}\)) and latent (\(\mathbf{x}_t\)). For example, see if it's possible to do collapsed Gibbs sampler? Or switch to EM as in PLDS? If I switch to EM, then the DPMM should also be done in EM. (That would be more challenging). \\

	Basically the question is
	\[y_{\text{it}} \sim Poisson(\lambda_{\text{it}})\]
	\[\log\lambda_{\text{it}} = d_{i} + \mathbf{c'}_{i}\mathbf{x}_{t}^{(z_{i})}\]
	\[\mathbf{x}_{1} \sim N(\mathbf{x}_{0},\mathbf{Q}_{0})\]
	\[\mathbf{x}_{t + 1}|\mathbf{x}_{t} \sim N(\mathbf{A}\mathbf{x}_{t} + \mathbf{b},\mathbf{Q})\]
	And we need to sample \(\mathbf{x}_t\), \(\mathbf{b}\), \(\mathbf{A}\) and \(\mathbf{Q}\) efficiently. (Assume loading is known now).

	Or even a more basic question: how to efficiently sample state-space and transition matrix at the same time in Kalman filter? The dependency between transition matrix and state vectors causes bad mixing.

	\item
	Check sampling trace and mixing.
	\item
	Find a more efficient way to generate new cluster parameters, otherwise the newly generated ones will always be rejected.
	\item
	improve DPMM. In current brute force implementation, the number of potential clusters can even go beyond number of neurons (\(N\)). There are several improvements, e.g. \href{https://link.springer.com/article/10.1007/s11222-009-9150-y}{Kalli et al., 2011}, \href{http://proceedings.mlr.press/v37/gea15.html}{Ge et al., 2015} and \href{https://link.springer.com/article/10.1007/s11222-014-9471-3}{Hastie et al, 2015}. Check them later.
\end{enumerate}

>>>>

The conditional priors for these parameters:
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	Latent vectors \(\left\{ \mathbf{x}_{t} \right\}_{t=1}^T\): the conditional prior is
	defined by
	\[\mathbf{x}_{1} \sim N(\mathbf{x}_{0},\mathbf{Q}_{0})\]
	\[\mathbf{x}_{t + 1}|\mathbf{x}_{t} \sim N(\mathbf{A}\mathbf{x}_{t} + \mathbf{b},\mathbf{Q})\]
	\item
	Initials \(\mathbf{x}_{0}\): assume there are \(J\) clusters,
	\[\mathbf{x}_{0} \sim N(\bm{\mu}_{\mathbf{x}_{00}},\ \mathbf{\Sigma}_{\mathbf{x}_{00}})\]
	, where \(\bm{\mu}_{\mathbf{x}_{00}} = \mathbf{0}_{\text{Jp}}\) and \(\mathbf{\Sigma}_{\mathbf{x}_{00}} = \mathbf{I}_{\text{Jp}}\)
	\item
	Linear mapping (loading) for latent vectors

	\textcolor{red}{\textbf{I just ignore that it is just multivariate GLM problem. Do it later...}}\\

	\item
	Mean and covariance for loading in each cluster

	\textcolor{red}{\textbf{MV-GLM...}}\\

	\item
	Linear dynamics for latent vectors \(\mathbf{A}\) and \(\mathbf{b}\): if the number of cluster is \(J\)

	\textcolor{red}{\textbf{MV-GLM...}}\\

	\item
	Process noise \(\mathbf{Q}\): if the number of cluster is \(J\)

	\textcolor{red}{\textbf{MV-GLM...}}\\

\end{enumerate}

@@

\subsubsection{Update \(\left\{ d_{i} \right\}_{i = 1}^{N}\) and
	\(\left\{ \mathbf{c}_{i} \right\}_{i = 1}^{N}\)}
To update efficiently, use Laplace approximation again. Denote

\textcolor{red}{\textbf{MV-GLM...}}\\

\subsubsection{Update \(\left\{ \mathbf{\mu}_{\text{dc}}^{(j)} \right\}_{j}\) and \(\left\{ \mathbf{\Sigma}_{\text{dc}}^{(j)} \right\}_{j}\)} \label{loading prior}

\textcolor{red}{\textbf{MV-GLM...}}\\

@@

\textcolor{red}{\textbf{MV-GLM...}}\\

\subsubsection{Update \(\mathbf{Q}\)}

\textcolor{red}{\textbf{MV-GLM...}}\\

\section{Simulations}

\textcolor{red}{\textbf{Old results}}\\

\textcolor{red}{\textbf{Currently, the latent \(x_t\) estimation is not accurate enough. This may come from bad simulation/ slow convergence or other sources. See details in section \ref{problems}}}\\

@@

\clearpage
\section{TODO}
\begin{enumerate}
	\def\labelenumi{(\arabic{enumi})}
	\item
	\textcolor{red}{\textbf{Do MV-GLM}}\\
	\item
	Check sampling trace and mixing.
	\item
	Find a more efficient way to generate new cluster parameters, otherwise the newly generated ones will always be rejected.
	\item
	improve DPMM. In current brute force implementation, the number of potential clusters can even go beyond number of neurons (\(N\)). There are several improvements, e.g. \href{https://link.springer.com/article/10.1007/s11222-009-9150-y}{Kalli et al., 2011}, \href{http://proceedings.mlr.press/v37/gea15.html}{Ge et al., 2015} and \href{https://link.springer.com/article/10.1007/s11222-014-9471-3}{Hastie et al, 2015}. Check them later.
	\item
	After all of them are resolved, switch to GP version

\end{enumerate}
